---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The student's t test

Now let's look at an example on how one would use the t test. We want to investigate whether or not there is a difference in weight between between male and female Larus argentatus. First, download the dataset **Larus argentatus.xlsx** from **[https://github repository not yet added]**. For this analysis, we'll use the two packages **readxl** and **ggpubr**. If this is the first time using these packages, you can install them using

```
install.packages("readxl")
install.packages("ggplot2")
```

``` {r}
library(readxl)
library(ggplot2)
```

Next, let's import the data. It can be a good idea to inspect the data to get an overview of the variables and the type of data we're dealing with. Using the `head()` command allows us to see the first rows of the dataset;

```` {r}
Data <- read_excel("Larus argentatus.xlsx")
head(Data, 10)
````

From the output we can see that the dataset contains the variables Sex, Location, Mass, and Date. Our research question was whether or not there is a difference in weight between males and females. Let's first look at the `Mass` means for both males and females;

```` {r}
mean(Data$Mass[Data$Sex == "M"])
mean(Data$Mass[Data$Sex == "F"])
````

We can also create a boxplot using **ggplot** to visualize the results;

```` {r}
ggplot(Data, aes(x = Sex, y = Mass)) +
  geom_boxplot() +
  xlab("") +
  ylab("Mass [g]") +
  scale_x_discrete(labels = c("Female", "Male")) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
````

It certainly looks like there is a difference in biomass. However, before conducting a t test to verify our hypothesis we need to check that our statistical assumptions are met. First, our variable of interest in the two groups have to follow a **normal distribution**. To check for normality in our variables, we can use the **Shapiro-Wilk test**. The input format in R is given by 

````
shapiro.test(dataset$variable[dataset$categorical predictor])
````
In our case, this would translate to

```` {r}
shapiro.test(Data$Mass[Data$Sex == "M"]) 
shapiro.test(Data$Mass[Data$Sex == "F"])
````
Both tests indicate a non-significant p-value (p > 0.05). This means that our distributions are not significantly different from a normal distribution so we can assume normality! Next, we need to check for **equal variance** (also known as homoscedasticity). In R, the input command is given by

````
var.test(Variable ~ Categorical predictor)
````
or, in our case;

````{r}
var.test(Data$Mass ~ Data$Sex)
````

The Fisher's test yields a non-significant p-value (p > 0.05), which implies that there's no difference in the variance between the two groups. This means that all our statistical assumptions for conducting a t test are met and we are greenlit to do a t test;

````{r}
t.test(Data$Mass~ Data$Sex)
````

From the output we see that the p-value is significant (p < 0.05), implying that there is indeed a statistical significant difference in biomass between male and female Larus argentatus.








